\chapter{Glossary} \label{chr:glossary}

\begin{description}
    \item[API] - Application Programming Interface

    \item[Backend] - Server side of the application in a client-server architecture

    \item[BB] - Ball bearing

    Phantom used in W-L test.

    \item[CAX] - Center of the radiation field

    As stated in Pylinac documentation (v3.25) this term is a misnomer, but it is used due to backwards-compatibility reasons. To ensure consistency, this term is also used in this thesis. \cite{pylinac_cax}

    \item[CSS] - Cascading Style Sheets

    The styling language used to describe the appearance of HTML elements.

    \item[CT] - Computed Tomography

    \item[CV] - Computer Vision

    \item[DICOM] - Digital Imaging and Communications in Medicine

    Technical standards and file formats commonly used to store medical images.

    \item[DRF] - Django Rest Framework

    \item[EPID] - Electronic Portal Imaging Device

    \item[Frontend] - Client side of the application in a client-server architecture

    \item[HTML] - HyperText Markup Language

    The markup language used to define the structure of a web page.

    \item[HTTP] - Hypertext Transfer Protocol

    \item[JS] - JavaScript

    The scripting language used to define custom logic on a web page.

    \item[LA analysis] - Leaves alignment analysis

    \item[Linac] - Linear particle accelerator
    
    \item[MLC] - Multileaf Collimator
    
    Part of Linac that controls beam geometry.

    \item[ORM] - Object Relational Mapping

    Layer that connects object-oriented programming languages with databases.

    \item[QC] - Quality control

    Process of verifying the characteristics of a given product and relating the results to the specified requirements.

    \item[RT] - Radiation therapy

    \item[SE] - Structuring element

    A shape used in morphological operations.

    \item[SID] - Source-Image Distance

    The distance between the radiation beam source and the image receptor.

    \item[UI] - User Interface
    
    Part of the application exposed to its users, that provides a medium of communication between application and users.

    \item[W-L test] - Winston Lutz test
\end{description}
